---
title: 'vildehaye sensors tutorials: Calculate AGL (Altitude above Ground) by Pressure'
author: "sivan margalit"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
---
## credits
This code is a translation of Sivan Toledo's MATLAB code.  
Integrating with DEM based on Eitam Arnon's code from his ViewShed application.  
The data, contributed by Dr. Yoni Vortman and Anat Levi, of navigating study done in Hula valey at 2020

## Background
  After loading Temperature and pressure measurements from vildehaye sensors, you can calculate Altitude above Sea Level (AGL) by pressure data. 
For this calculation we need information about the environment barometric data on same day. We can download this information from the Israely Meteorological Service (IMS)
Accessing to IMS data requires a Token (access key).  [here a link](https://ims.gov.il/he/ObservationDataAPI) for more information.

  Next step can be trasform the ASL to AGL (Altitude above Ground Level). We use DEM raster (Digital Elevation Model) to map of ground level's altitude.
  
### project structure
<ol>
<li>Data Folder:
  <ul> 
  <li>sqlite file of animal's localizations, which loaded from ATLAS database.</li>
  <li>binary data from battery and barometric sensors.</li>
  <li>excel file of meteorology stations, downloaded from IMS</li> 
  <li>DEM folder, which contains DEM rasters</li></ul>
</li>
  
<li>Function folder:
 <ul> 
  <li>pressure.R : contains 3 local functions, used to calculate ASL by BME information.</li>
  <li>DEM.R : implements the functionality of transform ASL to AGL </li> 
 </ul></li>
<li>VHsensors_project.Rproj : R project file, define the folder, where it located, as a working folder for R project</li>
</ol>



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(always_allow_html=TRUE)
ims.token='f058958a-d8bd-47cc-95d7-7ecf98610e47'
```
## R Code sample 
### loading libraries and local sources
The package "toolsForAtlas" contains the methods for read binary sensors data, and access to remote database by url address.

in functions folder we contains local source files contains functions, which are not in the package

```{r load libraries, message = FALSE}
library(toolsForAtlas) # useful functions for atlas users
options("digits"=14)   # show long numbers till 14 digits
Sys.setlocale("LC_TIME","C")  # change local date to English names
source("functions/pressures.R", encoding = "UTF-8")
source("functions/DEM.R", encoding = "UTF-8")
```

### loading data from binary files
```{r extract_data_1, message=c(1:4)}
# loading battery data
bat_df<-read_BAT("data/tag-972001006611-BATMON.bin")
bat_df$TAG<-"972001006611"
head(bat_df, 4)

# loading barometric pressure data (in mbar units)
pressure_df<-read_BME("data/tag-972001006611-BME280.bin")
pressure_df$TAG<-"972001006611"
head(pressure_df, 4)
```

### plot baromeric data
The file functions/pressures.R contains "plot_BME" function plotting barometric data.
This function has 4 arguments :
1. bme.df - is the sensor's data
2. reference.df - optional . a data frame of environment  barometric data, as a reference. when this data frame is supply the function add it to the BME plot
3. fromTime - optional. if it not supply the fromTime is the minimum time in the dataframe
4. toTime - optional. if it not supply the toTime is the maximum time in the dataframe
```{r plot_bme_1}
bme.plot.1<-plot_BME(bme.df=pressure_df)
bme.plot.1
```

### read local meteorologic data
The file functions/pressures.R contains "get.ISM.data" function which download pressure and temperature data, from a close meteorologic station.  
In this example - it uses stattion in Har-Knaan, in Tzfat 
```{r read_IMS}
# ims.token<- Insert token from IMS, and uncomment this line
ism.lst<-get.ISM.data(easting=257000,
                      northing=780000,
                      fromTime=min(pressure_df$TIME),
                      toTime=max(pressure_df$TIME),
                      ims.token)

ism.pressure.df<-ism.lst[["pressure"]]
ism.temperatures.df<-ism.lst[["temeratures"]]
hsl<-ism.lst[["hsl"]]

print(sprintf("Station Heigh relative to sea level=%d meter",hsl))
head(ism.pressure.df)
head(ism.temperatures.df)
```
(If you get red warning from red.xlsx - it is harmless)
lets plot the barometric data, including the refence data, retreived from the local IMS station
```{r plot_bme_2}
bme.plot.2<-plot_BME(bme.df=pressure_df, 
                     reference.df = ism.pressure.df)
bme.plot.2
```

### Estimate Altitude above Sea Level (ASL) by pressure
the function sensorsPressureToAltitude() in functions/pressures.R calculate altitude, using the formula TBD
This function has 4 arguments :
<ol>
<li>pressure_df - is the sensor's data</li>
<li>ism.pressure.df - a data frame of environment  barometric data, as a reference.</li>
<li>refTemperature - temperature drymeasurements in IMS station.</li>
<li>refAltitude - IMS station high above sea level.</li>
</ol>
```{r PressureToAltitude}
ASL.df<-sensorsPressureToAltitude(pressure=pressure_df,
                                   refPressure=ism.pressure.df,
                                   refTemperature=ism.temperatures.df,
                                   refAltitude=hsl)
```

now - show the altitude plot
```{r plot_altitude}
ASL.df$dateTime<-as.POSIXct(ASL.df$TIME, tz="UTC", origin="1970-01-01")
ggplot(ASL.df, aes(x=dateTime, y=ASL))+
  labs(title="Altitude Above Sea Level",
       y="ASL (meters)")+
  geom_point(aes(colour = factor(TAG)), size=0.5)+
  theme_bw()+
  theme(legend.title = element_blank())
```
### integrade ASL with ATLAS TAG data  
now read localization data:
```{r load_atlas_data}
atlas.df<-loadFromSQLite(dbn="data/atlas-6611.sqlite")
glimpse(atlas.df)
```
dplyr::glimpse() function show each column type. 
We recommending the following changes :
<ul>
<li>If TIME field is int64 , than change TIME field to double - since int64 does not handle well with compering operators</li>
<li> delete redundant columns to compact the data frame</li>
<li> it is nice to add dateTime + Date, too</li>
<li> also - you can change TAG to character (or factor)</li>
</ul>

```{r organized_loc_data}
atlas.df<-atlas.df%>%
  dplyr::select(-c(TX, Z, DIM, VARZ, COVXZ, COVYZ))%>%
  mutate(TIME=as.double(TIME),
         TAG=as.character(TAG),
         dateTime=as.POSIXct(TIME/1000, tz="UTC",origin="1970-01-01"),
         date=as.Date(dateTime))
         
glimpse(atlas.df)
```
merge ASL data with ATLAS localizations data, by TIME
```{r ASL_to_atlas.df}
  addAtlasLoc <- merge(x=atlas.df%>%
                         mutate(mergeTIME=round(TIME/1000)),
                       y=ASL.df%>%
                         dplyr::select(-dateTime), 
                       by.x=c("mergeTIME","TAG"),
                       by.y=c("TIME","TAG"),
                       all.x=TRUE)
  
  head(addAtlasLoc)

```

### Estimate Altitude above Ground Level (AGL) by ASL
now, we use function getAGLFromDEM() which uses the DEM file(s) uploaded to data/DEM/  to calculate AGL
```{r ASL_to_AGL}
  # get atlitude above ground , using DEM rasters 
  AGL.lst<-getAGLFromDEM(addAtlasLoc)
  
  # the function getAGLFromDEM() return a list with 2 items
  names(AGL.lst)
  
  # extract alas.based data frame from the AGL.lst
  alt.above.ground.df<-AGL.lst[["atlas.df"]]
  
  
  #show the altitude plot
  ggplot(ASL.df, aes(x=dateTime, y=ASL))+ 
    geom_point(color="red", size=0.5)+
    geom_point(data=alt.above.ground.df,
               aes(x=dateTime, y=AGL), color="blue", size=0.5)+
    #scale_x_datetime(limits = as.POSIXct(c('2020-12-02 07:00','2020-12-02 16:00')))+
    labs(title = "Altitude Above Ground Level",
       y="AGL (meters)")+
    theme_bw()+
    theme(legend.title = element_blank())

  
  # show histogram of data's AGL
  ggplot(alt.above.ground.df,)+
    geom_histogram(aes(AGL))+
    labs(title = "histogram of AGL",
      x="AGL (meters)")+
    theme_bw()+
    theme(legend.title = element_blank())



```
lets show the AGL over the geographic map, with leaflet
```{r AGL_leaflet}
AGL_ll(alt.above.ground.df)

```


